\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}

\title{Homework\#8}
\author{Raja Kantheti}
\date{\today}

\begin{document}

\maketitle

\section{Sample Graph}

We will use the following graph for our examples:
\begin{itemize}
    \item Vertices: \( V = \{A, B, C, D\} \)
    \item Edges with weights: 
    \begin{itemize}
        \item \( (A, B, 1) \)
        \item \( (A, C, 3) \)
        \item \( (B, C, 3) \)
        \item \( (B, D, 6) \)
        \item \( (C, D, 4) \)
    \end{itemize}
\end{itemize}

\section{Prim's Algorithm with Weight Matrix and Unordered Array}

\subsection{Steps}

\begin{enumerate}
    \item \textbf{Initialization:}
    \begin{itemize}
        \item Start with vertex \( A \).
        \item Priority queue (unordered array): \( [(A, 0)] \)
        \item MST: \( \emptyset \)
    \end{itemize}
    \item \textbf{Main Loop (runs |V| times):}
    \begin{itemize}
        \item Extract minimum: \( (A, 0) \)
        \item Add edges \( (A, B, 1) \) and \( (A, C, 3) \) to the queue.
        \item Priority queue: \( [(B, 1), (C, 3)] \)
        \item MST: \( \{A\} \)
        \item Extract minimum: \( (B, 1) \)
        \item Add edge \( (B, D, 6) \) to the queue.
        \item Priority queue: \( [(C, 3), (D, 6)] \)
        \item MST: \( \{A, B\} \)
        \item Extract minimum: \( (C, 3) \)
        \item Add edge \( (C, D, 4) \) to the queue.
        \item Priority queue: \( [(D, 4)] \)
        \item MST: \( \{A, B, C\} \)
        \item Extract minimum: \( (D, 4) \)
        \item Priority queue: \( \emptyset \)
        \item MST: \( \{A, B, C, D\} \)
    \end{itemize}
\end{enumerate}

\section{Prim's Algorithm with Adjacency List and Min-Heap}

\subsection{Steps}

\begin{enumerate}
    \item \textbf{Initialization:}
    \begin{itemize}
        \item Start with vertex \( A \).
        \item Priority queue (min-heap): \( [(0, A)] \)
        \item MST: \( \emptyset \)
    \end{itemize}
    \item \textbf{Main Loop (runs |V| times):}
    \begin{itemize}
        \item Extract minimum: \( (0, A) \)
        \item Add edges \( (A, B, 1) \) and \( (A, C, 3) \) to the heap.
        \item Priority queue: \( [(1, B), (3, C)] \)
        \item MST: \( \{A\} \)
        \item Extract minimum: \( (1, B) \)
        \item Add edge \( (B, D, 6) \) to the heap.
        \item Priority queue: \( [(3, C), (6, D)] \)
        \item MST: \( \{A, B\} \)
        \item Extract minimum: \( (3, C) \)
        \item Add edge \( (C, D, 4) \) to the heap.
        \item Priority queue: \( [(4, D), (6, D)] \)
        \item MST: \( \{A, B, C\} \)
        \item Extract minimum: \( (4, D) \)
        \item Priority queue: \( [(6, D)] \)
        \item MST: \( \{A, B, C, D\} \)
    \end{itemize}
\end{enumerate}

\section{Kruskal's Algorithm with Adjacency List and Fast Sort}

\subsection{Steps}

\begin{enumerate}
    \item \textbf{Initialization:}
    \begin{itemize}
        \item Sort edges by weight.
        \item Edges: \( [(A, B, 1), (A, C, 3), (B, C, 3), (C, D, 4), (B, D, 6)] \)
        \item MST: \( \emptyset \)
    \end{itemize}
    \item \textbf{Main Loop (runs |E| times):}
    \begin{itemize}
        \item Edge \( (A, B, 1) \) - add to MST.
        \item MST: \( \{(A, B, 1)\} \)
        \item Edge \( (A, C, 3) \) - add to MST.
        \item MST: \( \{(A, B, 1), (A, C, 3)\} \)
        \item Edge \( (B, C, 3) \) - cycle detected, skip.
        \item Edge \( (C, D, 4) \) - add to MST.
        \item MST: \( \{(A, B, 1), (A, C, 3), (C, D, 4)\} \)
        \item Edge \( (B, D, 6) \) - cycle detected, skip.
    \end{itemize}
\end{enumerate}

\end{document}
